### 2025-04-18 research-methods.qmd
20250219 - causal inference
20250417 - attenuation of associations due to unreliability

In an *experiment*, there are one or more things (i.e., variables) that we manipulate to see how the manipulation influences the process of interest.
The variable that we manipulate is the *independent variable*.
By contrast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).
Besides the independent and dependent variables, the researcher attempts to hold everything else constant through processes including standardization and random assignment.
*Standardization* involves using the same procedures to assess each participant, so that scores can be fairly compared across participants (and groups).
Random assignment involves randomly assigning participants to conditions of the independent variable, so the people in each condition are comparable and do not differ systematically.

@@

For more information on these and other aspects of reliability, see <https://isaactpetersen.github.io/Principles-Psychological-Assessment/reliability.html> [@Petersen2024a; @PetersenPrinciplesPsychAssessment].

>>>>

In an *experiment*, there are one or more things (i.e., variables) that we manipulate to see how the manipulation influences the process of interest.
The variable that we manipulate is the *independent variable*.
By contrast, the *dependent variable* is the variable that we evaluate to determine whether it was influenced by the manipulation (i.e., by the independent variable).
Besides the independent and dependent variables, the researcher attempts to hold everything else constant through processes including standardization and random assignment.
*Standardization* involves using the same procedures to assess each participant, so that scores can be fairly compared across participants (and groups).
Random assignment involves randomly assigning participants to conditions of the independent variable, so the people in each condition are comparable and do not differ systematically.
However, not all things are feasible or ethical to manipulate.
For instance, it would not be ethical to randomly assign some players to receive a head trauma.
In addition, it would not be feasible to manipulate the weather of a game or the performance of players.

@@

For more information on these and other aspects of reliability, see <https://isaactpetersen.github.io/Principles-Psychological-Assessment/reliability.html> [@Petersen2024a; @PetersenPrinciplesPsychAssessment].
Reliability is important to consider because random measurement error (i.e., unreliability) weakens the associations between variables [@Goodwin2006; @Schmidt1996].
